---
title: "BUA 455 Lecture 10"
author: "Penelope Pooler Eisenbies"
date: "`r Sys.Date()`"
output: 
  html_document: 
    code_folding: show
    toc: yes
    toc_depth: 4
    toc_float: yes
---

### Getting Started with Markdown

- Download Zipped R project 

- Open Zipped folder and copy internal folder (R Project) to a BUA 345 folder on your computer NOT IN DOWLOADS

- Click on .Rproj file to open project

- Click on `code_data_output` file to open it.

- Click on `BUA_345_Lecture_10.Rmd` to open it.

- Run Setup Chunk.

---

### Setup

- The setup chunk shows the packages needed for this demo.   

- R will install specified packages if needed (only required once after R is installed)  

- R will load specified packaged (required every time you start a new R session)  

- The first time you run this code, R will install these packages which will be slow.  

- **If you get warnings, that's okay.**  

- If you get **error messages**, I (or TA), can help you.

---

### Setup Chunk for Lecture 10

```{r setup for Lecture 10, include = T}


# this line specifies options for default options for all R Chunks
knitr::opts_chunk$set(echo=T, highlight=T)

# suppress scientific notation
options(scipen=100)

# install helper package that loads and installs other packages, if needed
if (!require("pacman")) install.packages("pacman", repos = "http://lib.stat.cmu.edu/R/CRAN/")

# install and load required packages
pacman::p_load(pacman,tidyverse, magrittr, olsrr)

# verify packages
p_loaded()

```

---

### Lecture 10 In-class Exercises

#### **Question 2 (L10) - Session ID: bua345s23**
  
Below is small R Chunk with a calculation.
 - `#` before text indicates a comment. R will ignore this text.
 
```{r first calculations in R}
(10*3) + sqrt(987) - (2*4) + 4567               # unrounded calculation
((10*3) + sqrt(987) - (2*4) + 4567) |> round(2) # rounded to 2 decimal places

```

Experiment with running the lines one at a time and running th whole chunk.

The output should appear below the chunk and in the Console (Lower Left Pane)  

What is the value from these calculations, rounded to 2 decimal places?

A. 4620.12
  
B. 4680.42
  
C. 4660.32
  
D. 4620.42

---

### Lecture 10 In-class Exercises

#### **Question 3 (L10) - Session ID: bua345s23** 

#### Creating and Saving Objects in R - Entering small datasets manually 

**`1.`** Copy and paste the R code shown below into the provided R Chunk.

```{r summarizing ages data}
ages <- c(21,18,27,43,56,23,23,25,29,40)   # create ages data vector
median(ages)                               # calculate median of ages
summary(ages)                              # calculate summary statistics for ages data

```


**`2.`** Run this code to create a numerical variable named `ages`.

**`3.`** Notice that `ages` appears in the `Global Environment`.

**`4.`** Find the median of `ages` using the `median` command.

The median of the `ages` data is:
  
A. 24  
B. 25  
C. 26  
D. 27  
E. 28  

---


### Importing Data into R

- In BUA 345, all data files will be stored in R Project `code_data_output` folder with R Markdown file

- `tidyverse` package must be loaded in `setup` chunk above

- running this code will save `houses` dataset to the `Global Environment`.

- `#` before text indicates a comment. R will ignore this text.**

```{r import houses data}
houses <- read_csv("houses.csv")                     # import houses data
houses <- read_csv("houses.csv", show_col_types = F) # same command with message surpressed

# same command with glimpse afterwards to examine data
# pipe |> operator used to make code more efficient

houses <- read_csv("houses.csv", show_col_types = F) |> glimpse()                                                               
```

---

### Examining Data - Variable Summaries and Correlations in R
  

```{r variable summaries and correlations in R}

houses |> select(Living_Area) |> summary() # summary of living area (with piping)

summary(houses$Living_Area)                # summary of living area without piping

houses |> summary()                        # summary of houses data (with piping)

summary(houses)                            # summary without piping

houses |> cor()                            # correlation matrix for houses (with piping)

cor(houses)                                # correlation matrix for houses (without piping)

```

---

### Lecture 10 In-class Exercises

#### **Question 4 (L10) - Session ID: bua345s23** 

What is the maximum selling price in the houses data? 

Round answer to closest thousand (K)

A. 150 K     
B. 165 K    
C. 205 K    
D. 291 K    
E. 466 K    
F. 521 K    


```{r R code calculate rounded max values}
# unrounded using summary()
houses |> pull(Selling_Price) |> summary() 

 # unrounded using max()
houses |> pull(Selling_Price) |> max()     

# rounded to closest thousand
# round(-3) means round to third place before decimal (thousands)
houses |> pull(Selling_Price) |> summary() |> round(-3)
houses |> pull(Selling_Price) |> max() |> round(-3)

```

---

### Lecture 10 In-class Exercises

#### **Question 5 (L10) - Session ID: bua345s23** 

What is the correlation between `House_Age` and `Living_Area` in the `houses` dataset? Round answer to two decimal places. 

A. 0.77    
B. 0.71    
C. 0.66    
D. -0.22    
E. -0.38   
F. -0.52    
    
NOTE:  All values in correlation matrix can be rounded at once with `round` command.

```{r  calculate rounded correlation matrix}
houses |> cor() |> round(2)   # rounded correlation matrix to two decimal places with piping

```

---

### R code to create and view a MLR model


```{r slr houses model}
# slr model with just living area
(houses_slr <- ols_regress(Selling_Price ~ Living_Area, data = houses))

```

---

### Multiple Linear Regression (MLR) in R

- Extending this model to Multiple Linear Regression is straightforward.
- Additional terms are added to the model with a `+` sign between them:
- MLR model for houses data:


```{r mlr houses model}

# mlr model with living area, bathrooms and house age in model
(houses_mlr <- ols_regress(Selling_Price ~ Living_Area + Bathrooms + House_Age, data = houses))

```